---
title: "Example: Catch data only"
author: "Steven Martell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Catch Only}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


The catch-MSY method is suited for using only catch data to define the lower boundary of the underlying production required to sustain the observed catch data.

For this example, the Namimbian Hake data set will be used.

```{r}
	library(catchMSY)
	library(tidyr)
	library(ggplot2)
	library(GGally)
	library(mvtnorm)
	library(foreach)
	data(NamibianHake)
	names(hake)
	hake$data$index <- NA # to run in catchOnly mode
	head(hake$data)
```

The relative abundance index will be ignored for this example.  For a given sID object, the first step is to draw random samples for M, FMSY and MSY based on the prior info specified in the sID object.  This is accomplished with the `sample.sid` function where the second argument specifies the number of samples.  Next, use the `sir.sid` function to generate `code`s for each of the parameter samples.

```{r}
  hake$dfPriorInfo
	hake <- sample.sid(hake,500)
	hake <- sir.sid(hake)
	head(cbind(hake$S,hake$code))
```

Once these codes have been generated, its now possible to plot the corresponding distributions based on the Monte Carlo sample.

```{r,fig.show='hold'}
  .THEME <- theme_bw()
df2 <- with(hake,{
  df  <- ps.dt[code==0,]
	colnames(df) = data$year
	df <- cbind(id=1:dim(df)[1],df)
	return(gather(df,key,dt,-id))
})

p <- ggplot(df2,aes(x=as.double(key),y=dt)) 
p <- p + stat_summary(geom="ribbon",fun.ymin="min",fun.ymax="max",alpha=0.1)
p <- p + stat_summary(geom="ribbon", fun.data="median_hilow",alpha=0.2,fill="blue")
p <- p + stat_summary(geom="ribbon", fun.data="mean_cl_boot",alpha=0.2,fill="red")
p <- p + geom_line(aes(group=id),size=0.1,alpha=0.1)
p <- p + labs(x="Year",y="Depletion")
print(p + .THEME)

df<- with(hake,{
  df <- data.frame(S,code=code,clr='black')
	pdf <- subset(df,code==0)
	ndf <- subset(df,code!=0)
	ndf$clr <- 'red'
	return(rbind(pdf,ndf))
})

# ggpairs(df,1:3,color='clr')
p <- ggpairs(df,1:3,color='clr',upper = list(continuous = "density", combo = "box"))
print(p + .THEME)



```
