---
title: "Example: Namibian Hake"
author: "Steve Martell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{Namibian Hake Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


##Summary
In this vignette I use the Namimian Hake as an example of how to use this package using:

  1. only catch data
  2. catch data and relative abundance
  3. catch data, relative abundance, and changes in mean size of the catch

First, we'll have a look at the input data and the Stock ID object `sID`.  To create a new stock ID object we use the  function `new_sID`. The data for this example were taken from chapter 10 in the Ecological Detective (Hilborn and Mangel, 1997).

```{r}
	library(catchMSY)
	data(NamibianHake)
	names(hake)
	head(hake$data)
```

Given a valid `sID' object, you can quickly run the age-structured model and preview the results.

```{r}
	hake <- runModel(hake)
	str(hake)
```


```{r, echo=FALSE, results='asis'}
library(catchMSY)
data(NamibianHake)
knitr::kable(head(hake$data, 10))
```
### Input data and Stock ID


The stock ID object for this example is included in the `catchMSY` package, and the default values for `new_sID` were used to set up the initial stock ID object.  Here is the structure of that object.





The `catchMSY` package is intended to be used to determine MSY-based reference points.

The catchMSY package is based on the initial work for Martell and Froese (2012). "A simple method for estimating MSY from catch and resilience".

This is a simple example based on the Namibian hake dataset published in the ecoloigical detective (Hilborn and Mangle, 1997).

Steps:

- assemble a time-series data file for your stock.
- create a stockID object

## stockID object (sID)

Describe the contaents of an 'sID' object.


```{r}
library(catchMSY)
library(ggplot2)


# load the hake sID object
data(NamibianHake)

# run the age structured model.
hake <- runModel(hake)

# plot the biomass
qplot(hake$year,hake$bt,geom="line")


```

## References

Hilborn, R. and Mangel, M. (1997). The ecological detective: confronting models with data. Princeton Univ Pr.

Martell, S. and Froese, R. (2012). A simple method for estimating msy from catch and resilience. Fish and Fisheries.

